---
title: "Assigment - kNN DIY"
author:
  - Thanh Dung Nguyen - Author
  - Duyen Nguyen - Reviewer
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   html_notebook:
    toc: true
    toc_depth: 2
---

```{r}
library(tidyverse)
library(googlesheets4)
library(class)
library(caret)
library(magrittr)
library(gmodels)
```

------------------------------------------------------------------------

Choose a suitable dataset from [this](https://github.com/HAN-M3DM-Data-Mining/assignments/tree/master/datasets) folder and train your own kNN model. Follow all the steps from the CRISP-DM model.

## Business Understanding
Viruses in the blood of blood donors are the cause of a number of diseases. Machine learning (or KNN) can automate and improve the efficiency and accuracy of virus screening. The risk of infection through blood transfusion is also limited and controlled

## Read data
Read data in csv format from url

```{r}
url <- "https://raw.githubusercontent.com/HAN-M3DM-Data-Mining/assignments/master/datasets/KNN-hcvdat0.csv"
rawDF <- read_csv(url)
View(rawDF)
```

## Data Understanding
Get an overview of basic information about the dataset
The dataset has 14 variables (columns) and 615 observations (rows)
Variables Category and Sex have character data type, others have numeric 

```{r}
str(rawDF)
```

## Data Preparation
remove the first column from the dataset which contains irrelevant information and the NA values

```{r}
cleanDF <- rawDF[-1]
cleanDF1 <- drop_na(cleanDF)
cleanDF2 <- na.omit(cleanDF1)
head(cleanDF2)
```

#checking the counts and proportions of the outcomes (variable Category)
```{r}
cntCat <- table(cleanDF2$Category)
propCat <- round(prop.table(cntCat) * 100 , digits = 1)

cntCat
propCat
```

#setting factor type for the variable
```{r}
cleanDF2$Category <- factor(cleanDF2$Category, levels = c("0=Blood Donor","0s=suspect Blood Donor", "1=Hepatitis", "2=Fibrosis", "3=Cirrhosis"), labels = c("Blood Donor", "Cannot give blood", "Cannot give blood", "Cannot give blood", "Cannot give blood")) %>% relevel ("Cannot give blood")
head(cleanDF2, 10)
view(cleanDF2)
```

```{r}
summary(cleanDF2[c("ALB", "ALP", "ALT")])
```
#normalization function
```{r}
normalize <- function(x) { 
  return ((x - min(x)) / (max(x) - min(x))) 
}

testSet1 <- c(1:5)
testSet2 <- c(1:5) * 10

cat("testSet1:", testSet1, "\n")
cat("testSet2:", testSet2, "\n")
cat("Normalized testSet1:", normalize(testSet1), "\n")
cat("Normalized testSet2:", normalize(testSet2))
```
#normalizing data
```{r}
nCols <- dim(cleanDF2)[2]
cleanDF_n <- sapply(4:nCols,
                    function(x) {
  normalize(cleanDF2[, x])
}) %>% as.data.frame()

summary(cleanDF_n[c("ALB", "ALP", "ALT")])
```
#split data set into training and test sets
```{r}
set.seed(123)
data_spl<-sample(1:nrow(cleanDF_n),size=nrow(cleanDF_n)*0.8,replace = FALSE)
train<-cleanDF_n[data_spl,]
test<-cleanDF_n[-data_spl,]
```

#create separate training and tests sets for labels
```{r}
train_labels <- cleanDF2[data_spl, 1]
test_labels <- cleanDF2[-data_spl,1]
```



## Modeling
```{r}
cleanDF_test_pred <- knn(train = as.matrix(train), test = as.matrix(test), cl = as.matrix(train_labels), k = 11)
head(cleanDF_test_pred)
confusionMatrix(cleanDF_test_pred, test_labels[[1]], positive = NULL, dnn = c("Prediction", "True"))
```

## Evaluation and Deployment

```{r}
CrossTable(x=as.matrix(test_labels), y=cleanDF_test_pred, prop.chisq=FALSE)
```
The total accuracy of the model is 92%, no cases of False Positives (FP), 9 cases of False Negatives (FN), meaning 9 cases were actually Cannot give blood in nature but got predicted as Blood Donor. This is considered risky in the case of a medical model

